### 2016-02-03 chapters/abstract.tex
restructuring and changing for english style. starting to include a bibliography for references

Variability management is an essential part of working on product lines.
This paper considers existing methods of configuration and tries to come up with a better alternative based on questionnaires to enable users or customers to configure a product on their own and to allow experts to design the questionnaires according to their domain knowledge.
To further simplify the process of product configuration out of software product-lines, feature models are used to describe the set of features and constraints contained in a given software product-line. This paper proposes a method for automatically generating feature models out of descriptive files and naming conventions.