### 2016-02-03 chapters/conclusion.tex

\subsection{Conclusion}
This is where the work is concluded. In this section there will be a description of the way we did things and the experiences we made during it. An emphasis will be on the insights and the findings from the scenarios will get outlined.

\subsection{Future Work}
Here will be a summary of the new questions that were raised in this work. Also there will be topics for further research. Particularly the problems we encountered and couldn't solve with our concept and why will be pointed out and first approaches will be suggested.