### 2016-02-12 chapters/abstract.tex
short notes for each section

Variability management is an essential part of working on product lines.
As an established way to simplify the process of product configuration out of software product-lines, feature models are used to describe the set of features and constraints contained in a given software product-line. This paper proposes a method for automatically generating feature models out of descriptive files and naming conventions.
Furthermore existing methods of configuration are considered in this work and to develop an alternative based on questionnaires to enable users or customers to configure a product on their own and to allow experts to design the questionnaires according to their domain knowledge.