### 2016-03-26 chapters/scenarios.tex

State the process of FM-Extraction on the example of \textit{Odoo}.

We applied these steps to a total of [A] files, thus extracting [B] features for \textit{Odoo}. These features are arranged in a total of [C] hierarchical levels. Additionally, [D] constraints limit the possible valid configurations of \textit{Odoo}.

Also state how some of the questions and their mapping in the questionnaire were developed.

Finally, show the process of configuration through running a fictinal user over the questionnaire and document his decisions as well as the resulting configurations.

%TODO: test ausdenken, der alle ``Fähgkeiten'' zeigt

% Fahrradladen, Online-Shop, Social-Media, Bezahldienste, Kontakt-Portal, DE/EN, Newsletter
%
% Website builder, eCommerce, SEA
% Sales, CRM, invoicing, Point of Sale
% Inventory, Purchase
% Email Marketing, Survey

%TODO: Planung, durchführung, Auswertung

%TODO: nur ein konstruierter Fall, weil Domain Knowledge fehlt.