### 2016-03-31 chapters/abstract.tex
rewrote the abstract

Variability management is an essential part of working on product lines. Feature models are an established tool to describe the set of features and constraints contained in a given software product-line, thus showing the variability. Feature models also facilitate configuration of product lines. However, for most software product lines there exists no feature model. This paper proposes a method for automatically generating feature models out of the structure of given source code and naming conventions.

Furthermore, an alternative method of configuration based on questionnaires is developed in this work. It enables users or customers to configure a product on their own. Also, it allows experts to design the questionnaires according to their domain knowledge during implementation of the product line.